---
title: "WineProject"
author: "Petra Yousefi, Sadaf fakhriravri, and Ani"
date: "12/2/2020"
output:
  pdf_document: default
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#import libraries
```{r}
library(car)
library(alr4)
library(leaps)
library(mice)
library(dplyr)
library(corrplot)
library(ggplot2)
library(visreg)
library(ggcorrplot)
library(corrr)
library(VIM)
library(moments)
library(tidyverse)
library(caret)
library(MASS)
library(leaps)
library(Simpsons)
library(memisc)
library(caret)
library(ggplot2)
```
#Lets first read out data file
#we are looking at red and white wines and what factors impact their quality.

```{r}
df<-read.csv("/Users/jamesdraper/Documents/R/redwine.csv",header=TRUE)
str(df)
```
#information about our data
-We have got 13 variables(columns)(attributes)\
-6497 observations\
-fixed acidity: numerical 
-volatile acidity: numerical
-citric acidity: numerical
-residual sugar: numerical
-chlorides: numerical
-free sulfur dioxide:numerical
-total sulfar dioxide:numerical
-density:numerical
-ph: numerical
-sulphates:numerical
-alcohol:numerical
-quality:int
-type:char
```{r}
summary(df)
as.double(df$quality)
df=na.omit(df)
str(df)
```
we can conclude there is no missing data within our data set. 
we want to check for outliers from one or more data
using Zscore of 3 .1% of population
we are going to do this not using zscore because that is the distance from the mean we want the distance from the centriod mean of means
mahalanobis^
using chi square 
p<0.001
degrees of freeom is going to be the number of variables to calculate the mahalanobis
once i find the outliers im going to delete them since im finding such a little amount \
leverage values and influence values later one 

```{r}
#getting rid of outliers for model
head(df)
dfNoRed<-df[ ,-c(13)]
head(dfNoRed)
mahal1<-mahalanobis(dfNoRed,
                    colMeans(dfNoRed,na.rm=TRUE),
                    cov(dfNoRed, use='pairwise.complete.obs'))
summary(mahal1)
cutoff=qchisq(.999,ncol(dfNoRed))
summary(mahal1<cutoff)
NoOut<-df[mahal1<cutoff, ]
summary(NoOut)
str(NoOut)
View(NoOut)
dfnew<-NoOut[ ,-c(13)]
View(dfnew)
```

#now to check for mulitcolinearity or singularity
so we are going to check with a correlation table.
so once we look at the scores if one or more have too high of a score we need to pick one or combine both (in my case im going to pick between them)
```{r}
corelation<-cor(dfNoRed,use='pairwise.complete.obs' )
symnum(corelation)
#we are good lol 
```
```{r}
#creating histograms too see which variables have skewness or need transformation
fa=NoOut$fixed.acidity
va=NoOut$volatile.acidity
ca=NoOut$citric.acid
rs=NoOut$residual.sugar
ch=NoOut$chlorides
fsd=NoOut$free.sulfur.dioxide
tsd=NoOut$total.sulfur.dioxide
d=NoOut$density
ph=NoOut$pH
s=NoOut$sulphates
a=NoOut$alcohol
q=NoOut$quality
hist(fa)
hist(va)
hist(ca)
hist(rs)
hist(ch)
hist(fsd)
hist(tsd)
hist(d)
hist(ph)
hist(s)
hist(a)
hist(q)
```
From the histograms we can tell that there is skeweness in alcohol, sulphate, total sulfur dioxide, free sulfur dioxide, chloride, residual sugars, citric acid, and volatile acid, and fixed acidity. Basically everything expect for density and ph has skeweness.

#now to check for linearity
```{r}
m1<-lm(q~fa+va+ca+rs+ch+fsd+tsd+d+ph+s+a)
summary(m1)
#checking again to see what is important to the model or not
regfit.full=regsubsets(quality~.,dfnew)
summary(regfit.full)
regfit.full=regsubsets(quality~., data=dfnew, nvmax=11)
reg.summary=summary(regfit.full)
names(reg.summary)
reg.summary$cp
vif(m1)
par(mfrow=c(1,2))
#boxcox(m1,lambda=seq(0,1,length=20))
#boxcox(m1,lambda=seq(0,1,length=20))
ppois(q= 4, lambda = 13, lower.tail = TRUE)
dpois(x= 0:4, lambda = 13)

```
```{r echo=FALSE, message=FALSE, warning=FALSE}
#1
ggplot(data = dfnew, aes(x = quality, y = fixed.acidity)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
#we can see that fixed acidity has almost no effect on Quality
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
#2
ggplot(data=dfnew, aes(x = quality, y = volatile.acidity)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
#volatile acidity has a negative affect on the quality where if VA goes up quality goes down
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
#3
ggplot(data=dfnew, aes(x=quality, y=citric.acid)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
#citric acid seems to have a positive affect higher quality higher citric acid
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
#4
ggplot(data=dfnew, aes(x=quality, y=residual.sugar)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = 'blue') +
  scale_y_continuous(lim = c(0,5)) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
#residual sugar seems to have little to no affect on quality (sweetness)
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
#5
ggplot(data=dfnew, aes(x=quality, y=chlorides)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = 'blue') +
  scale_y_continuous(lim = c(0,0.2)) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
#chlorides seems to have very little affect on quality but the less the chlorides a little better quality (salt)
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
#6
ggplot(data=dfnew, aes(x=quality, y=free.sulfur.dioxide)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = 'blue') +
  scale_y_continuous(lim = c(0,40)) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
#interesting... low FSD low quality however higher FSD average quality
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
#7
ggplot(data=dfnew, aes(x=quality, y=total.sulfur.dioxide)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = 'blue') +
  scale_y_continuous(lim = c(0,150)) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
#same thing here with total sulfur dioxide
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
#8
ggplot(data=dfnew, aes(x=quality, y=density)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
#better wines seem to have lower densities however could be from alcohol since low density means higher alcohol which is an important factor for wine.
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
#9
ggplot(data=dfnew, aes(x=quality, y=pH)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
#lower pH better wine (more acidic) (FURTHER ANALYSIS)
ggplot(data = dfnew, aes(x = fixed.acidity, y = pH)) +
  geom_point(alpha = 0.3) +
  scale_x_log10(breaks=seq(5,15,1)) +
  xlab("Fixed Acidity in Log Scale") +
  geom_smooth(method="lm")
ggplot(data = dfnew, aes(x = volatile.acidity, y = pH)) +
  geom_point(alpha = 0.3) +
  scale_x_log10(breaks=seq(.1,1,.1)) +
  xlab("Volatile Acidity in Log Scale") +
  geom_smooth(method="lm")
ggplot(data = subset(dfnew, citric.acid > 0), aes(x = citric.acid, y = pH)) +
  geom_point(alpha = 0.3) +
  scale_x_log10() +
  xlab("Citric Acid in Log Scale") +
  geom_smooth(method="lm")

simpsons <- Simpsons(volatile.acidity, pH, data=dfnew)
plot(simpsons)

```

```{r, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)

#10
ggplot (data = dfnew, aes(x=quality, y=sulphates)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
#generally seems better quality have more sulphates
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
#11

ggplot(data=dfnew, aes(x=quality, y=alcohol)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
#higher alcohol seems to have an affect on quality but we can do a simple linear regression to see if big impact is made
ma<-lm(quality~alcohol, data=dfnew)
summary(ma)
#with my cleaned data we see that alcohol alone 25% of the wines quality so there has to be more to it 
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
simple_cor_test <- function(x, y) {
  return(cor.test(x, as.numeric(y))$estimate)
}
cor(dfnew)

correlations <- c(
  simple_cor_test(dfnew$fixed.acidity, dfnew$quality),
  simple_cor_test(dfnew$volatile.acidity, dfnew$quality),
  simple_cor_test(dfnew$citric.acid, dfnew$quality),
  simple_cor_test(log10(dfnew$residual.sugar), dfnew$quality),
  simple_cor_test(log10(dfnew$chlorides), dfnew$quality),
  simple_cor_test(dfnew$free.sulfur.dioxide, dfnew$quality),
  simple_cor_test(dfnew$total.sulfur.dioxide, dfnew$quality),
  simple_cor_test(dfnew$density, dfnew$quality),
  simple_cor_test(dfnew$pH, dfnew$quality),
  simple_cor_test(log10(dfnew$sulphates), dfnew$quality),
  simple_cor_test(dfnew$alcohol, dfnew$quality))
names(correlations) <- c('fixed.acidity', 'volatile.acidity', 'citric.acid',
                         'log10.residual.sugar',
                         'log10.chlordies', 'free.sulfur.dioxide',
                         'total.sulfur.dioxide', 'density', 'pH',
                         'log10.sulphates', 'alcohol')

correlations
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
fitAll <- lm(quality ~ ., data = dfnew)
fitStart <- lm(quality ~ 1, data = dfnew)
head(dfnew)
summary(fitStart)
step(fitStart, 
     direction = "forward", 
     scope = formula(fitAll))
```
```{r}
#after cross validation lets do partial F testing on reduced model and full model. 
#havent decided how im going to do the f-test yet 
```
```{r}
#cross validation
k=5

set.seed(1)
folds=sample(1:k,nrow(dfnew),replace=TRUE)
cv.errors=matrix(NA,k,11, dimnames=list(NULL, paste(1:11)))
for(j in 1:k){
  best.fit=regsubsets(quality~.,data=dfnew[folds!=j,],nvmax=11)
  for(i in 1:11){
    pred=predict(best.fit,dfnew[folds==j,],id=i)
    cv.errors[j,i]=mean( (dfnew$quality[folds==j]-pred)^2)
  }
}
mean.cv.errors=apply(cv.errors,2,mean)
mean.cv.errors
par(mfrow=c(1,1))
plot(mean.cv.errors,type='b')
reg.best=regsubsets(quality~.,data=dfnew, nvmax=11)
coef(reg.best,11)
```




```{r}
#linear model with variables from forward stepwise selection
#just for the sake to see if there is a difference
#and with the ideas of how alcohol correlates with the other predictors
m4<-lm(quality~log(alcohol)+log(volatile.acidity)+log(sulphates)+log(total.sulfur.dioxide)+pH+log(free.sulfur.dioxide)+citric.acid, data=dfnew)
summary(m4)
plot(m4)
#generalized linear model we have not covered in another course (learn by yourself)
#poisson?
```


```{r}
#after stepwise selection forward and also taking into consideration my skewed data
#so i can use a log function for a transformation of my data
m3<-lm(quality~density*alcohol+alcohol*sulphates+(alcohol)+(volatile.acidity)+(sulphates)+(total.sulfur.dioxide)+pH+(free.sulfur.dioxide)+(citric.acid), data=dfnew)
summary(m3)
```
```{r}
#attempt number 3 of a linear model creating a training and testing set
set.seed(1534)
training_data <- sample_frac(dfnew, .6)
test_data <- dfnew[ !dfnew$X %in% training_data$X, ]
m1 <- lm(as.numeric(quality) ~ alcohol, data = training_data)
m2 <- update(m1, ~ . + log(sulphates))
m3 <- update(m2, ~ . + volatile.acidity)
m4 <- update(m3, ~ . + citric.acid)
m5 <- update(m4, ~ . + fixed.acidity)
m6 <- update(m2, ~ . + pH)
mtable(m1,m2,m3,m4,m5,m6)


```

#conclusion
So basically what I am thinking is that this data set has to many values where the qualities are around the average quality and therefore cannot create a better linear model

Created many different kinds of linear models however always get around the same results with an R^2 score of around .38. 



